
```{r}
library(ggplot2)
library(RMySQL)
library(edgeR)
library(WGCNA)
library(GGally)
library(DESeq2)
library(reshape)
library(flashClust)
```


```{r}

```{r open_connection}
#open SQL connection
if (exists("mydb")) dbDisconnect(mydb)
  mydb = dbConnect(MySQL(), dbname='mucrosquamatus-networks')
```


```{r ercc}
ercc <-dbGetQuery(mydb,'SELECT fpkm.*, mix1.mix1, mix2.mix2 FROM fpkm 
JOIN (SELECT id, attamoles_ul AS mix1 FROM ExFold WHERE mix=1) AS mix1
ON gene_id = mix1.id
JOIN (SELECT id, attamoles_ul AS mix2 FROM ExFold WHERE mix=2) AS mix2
ON gene_id = mix2.id')

factors <-dbGetQuery(mydb,'SELECT * FROM factors')
rownames(factors) <- factors$id
factors$days <- factor(factors$days)
factors$sex <- factor(factors$sex)

mix1<-melt(ercc[,c("gene_id","mix1",factors[factors$mix==1,"id"])],id=c("gene_id","mix1"))
mix2<-melt(ercc[,c("gene_id","mix2",factors[factors$mix==2,"id"])],id=c("gene_id","mix2"))
ggplot(mix2,aes(log(mix2,2),log(value,2),color=variable))+geom_point()
ggplot(mix1,aes(log(mix1,2),log(value,2),color=variable))+geom_point()

```

```{r dge}
fpkm <-dbGetQuery(mydb,'SELECT * FROM fpkm WHERE NOT gene_id REGEXP "^ERCC"')
rownames(fpkm) <- fpkm$gene_id  ; fpkm <- fpkm[,-1]
counts <-dbGetQuery(mydb,'SELECT * FROM counts WHERE NOT gene_id REGEXP "^ERCC"')
rownames(counts) <- counts$gene_id  ; counts <- counts[,-1]

keep <- rowMeans(fpkm)>=1  #eliminate genes with average fpkm >= 1
table(keep)
#design matrix

design <- model.matrix(~factor(days)+factor(sex)+0,factors)
rownames(design) <- rownames(factors)

#glm
dge <- DGEList(counts[keep,rownames(factors)])
dge <- calcNormFactors(dge)
dat <- estimateGLMCommonDisp(dge, design)
dat <- estimateGLMTrendedDisp(dat, design)
dat <- estimateGLMTagwiseDisp(dat, design)
fit <- glmFit(dat,design)

mds <- plotMDS(dge)
ggpairs(factors[,c("svLength", "tailLength", "totLength", "headLength",  "mass")])
mdsPlot <- data.frame(mds1=mds$cmdscale.out[,1], mds2=mds$cmdscale.out[,2], ID=rownames(mds$cmdscale.out),days=factors$days,sex=factors$sex,mass=factors$mass)
ggplot(mdsPlot,aes(mds1,mds2,shape=sex,color=factor(days),size=log(mass)))+geom_point()+my_theme()

oneVs8 <- glmLRT(fit,contrast=c(1,0,0,-1,0))  
summary(caste.de <- decideTestsDGE(oneVs8, p=0.05, adjust="BH"))
twoVs8 <- glmLRT(fit,contrast=c(0,1,0,-1,0))  
summary(caste.de <- decideTestsDGE(twoVs8, p=0.05, adjust="BH"))
fourVs8 <- glmLRT(fit,contrast=c(0,0,1,-1,0))  
summary(caste.de <- decideTestsDGE(fourVs8, p=0.05, adjust="BH"))
maleVsfemale <- glmLRT(fit,coef=5)  
summary(caste.de <- decideTestsDGE(maleVsfemale, p=0.05, adjust="BH"))

```


## Filter invariant probes, and plot scale free topology

```{r wgcna_soft_threshold, cache=TRUE}

variance_filter <- function(data) {
  # take WGNCA-style expression as input, and filter it for non-zero variance
  variancedatExpr=as.vector(apply(as.matrix(data),2,var, na.rm=T))
  return(data[, variancedatExpr>0 ])
}

soft_threshold <- function(data) {
  disableWGCNAThreads()
  # Choose a set of soft-thresholding powers, given a WGCNA data object
  powers = c(c(1:10), seq(from = 12, to=16, by=2))
  # Call the network topology analysis function
  sft = pickSoftThreshold(data, powerVector = powers, verbose = 5,networkType = "signed")
  # Plot the results:
  sizeGrWindow(9, 5)
  par(mfrow = c(1,2));
  cex1 = 0.9;
  # Scale-free topology fit index as a function of the soft-thresholding power
  plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
       xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
      main = paste("Scale independence"));
  text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
      labels=powers,cex=cex1,col="red");
  # this line corresponds to using an R^2 cut-off of h
  abline(h=0.90,col="red")
  # Mean connectivity as a function of the soft-thresholding power
  plot(sft$fitIndices[,1], sft$fitIndices[,5],
      xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
      main = paste("Mean connectivity"))
  text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
  par(mfrow = c(1,1));
    
}

# filter more to speed up WGCNA
keep2 <- rowMeans(fpkm)>=10
table(keep2) 
se <- DESeqDataSetFromMatrix(round(fpkm[keep2,]),factors[,c("days","sex")],formula(~days+sex))
vst <-varianceStabilizingTransformation(se)
vst <- rlog(se,fitType='local')
# filter data and plot soft threshold parameters
soft_threshold(variance_filter(t(assay(vst))))

```

Choosing power 12, since larger values tend to produce more problematic TOMs

```{r WGCNA}
enableWGCNAThreads()
wgcna_adjacency <- function(datExpr, threshold, minModuleSize=30, MEDissThres = 0.2, deepSplit = 2) {
  # takes WGNCA data expression matrix, a soft threshold
  # optionally minimum module size, module eigengene dissimilarity threshold, and splitting threthold
  # returns adjacency matrix, module eigengenes and a data frame of colors 
  
  #compute adjacency
  adjacency <- adjacency(datExpr, power = threshold,type="signed")
#  TOM <- TOMsimilarityFromExpr(datExpr,power = threshold, networkType="signed")
  TOM <- TOMsimilarity(adjacency,TOMType="signed")
  
  geneTree <- flashClust(as.dist(1-TOM), method = "average");
  # Module identification using dynamic tree cut:
  dynamicMods <- cutreeDynamic(dendro = geneTree, distM = 1-TOM, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize);
  table(dynamicMods)
  dynamicColors = labels2colors(dynamicMods)
  
  # Calculate eigengenes
  MEList = moduleEigengenes(datExpr, colors = dynamicColors)
  MEs = MEList$eigengenes
  # Calculate dissimilarity of module eigengenes
  METree = flashClust(as.dist(1-cor(MEs)), method = "average");
  plot(METree, main = "Clustering of module eigengenes",xlab = "", sub = "")
  # Plot the cut line into the dendrogram
  abline(h=MEDissThres, col = "red")
  
  merge <- mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 0)
  
  # The merged module colors
  mergedColors = merge$colors
  # Eigengenes of the new merged modules:
  mergedMEs = merge$newMEs
  
  # Rename to moduleColors
  moduleColors = mergedColors
  # Construct numerical labels corresponding to the colors
  colorOrder = c("grey", standardColors(50));
  moduleLabels = match(moduleColors, colorOrder)-1;
  MEs = mergedMEs;
  
  # Recalculate MEs with color labels
  invisible(MEs0 <- moduleEigengenes(datExpr, moduleColors)$eigengenes)
  MEs = orderMEs(MEs0)
  
  print(table(moduleColors))
  moduleColors <- as.data.frame(moduleColors)
  rownames(moduleColors) <- colnames(datExpr)
  
  return(list(adjacency=adjacency,MEs=MEs,moduleColors=moduleColors))
}
```

```{r}
datExpr<-t(assay(vst))
plotClusterTreeSamples(datExpr=datExpr)
results <- wgcna_adjacency(datExpr,12,MEDissThres = 0.2)
#results[["adjacency"]]
#results[["MEs"]]
table(results[["moduleColors"]])
results[["moduleColors"]]["c22249_g1",]
results[["moduleColors"]]["c65172_g1",]
results[["moduleColors"]]["c22290_g4",]
temp<-t(rbind(as.numeric(factors[colnames(fpkm),"days"]),fpkm["c22290_g4",]))
plot(temp[,1],temp[,2])
cor.test(temp[,1],temp[,2])
```


```{r module_trait}
plotModuleTrait <-function(moduleTraitCor,moduleTraitPvalue,species) {
  # plot heatmap with module-trait correlation data
  textMatrix = paste(signif(moduleTraitCor, 2), " (",
  signif(moduleTraitPvalue, 1), ")", sep = "");
  dim(textMatrix) = dim(moduleTraitCor)
  par(mar = c(6, 8.5, 3, 3));
  # Display the correlation values within a heatmap plot
  labeledHeatmap(Matrix = moduleTraitCor,
  xLabels = colnames(moduleTraitCor),
  yLabels = rownames(moduleTraitCor),
  colorLabels = FALSE,
  colors = blueWhiteRed(50),
  textMatrix = textMatrix,
  setStdMargins = FALSE,
  cex.text = .5,
  zlim = c(-1,1),
  main = paste("Module-trait relationships",species))
}

moduleTraitCor = cor(results[["MEs"]], as.numeric(factors$days), use = "p",method="spearman")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples=30)
print(table(moduleTraitPvalue < 0.05))

ggplot(data.frame(cor=moduleTraitCor,name=rownames(moduleTraitCor),pval=moduleTraitPvalue),aes(x=1,y=name,label=paste(signif(moduleTraitCor, 2), " (", signif(moduleTraitPvalue, 2), ")", sep = "")))+ geom_tile(aes(fill=cor))+ scale_fill_gradient(low = "steelblue",high = "#FF4D4D")+theme_bw()+geom_text()+xlab("")+ylab("")+theme(axis.text.x  = element_blank(),axis.ticks.x  = element_blank())

```
