
```{r}
library(ggplot2)
library(RMySQL)
library(edgeR)
library(WGCNA)
library(GGally)
library(DESeq2)
library(reshape)
library(flashClust)
library(GOstats)
library(GSEABase)
```


```{r}

```{r open_connection}
#open SQL connection
if (exists("mydb")) dbDisconnect(mydb)
  mydb = dbConnect(MySQL(), dbname='mucrosquamatus-networks')
```


```{r ercc}
ercc <-dbGetQuery(mydb,'SELECT fpkm.*, mix1.mix1, mix2.mix2 FROM fpkm 
JOIN (SELECT id, attamoles_ul AS mix1 FROM ExFold WHERE mix=1) AS mix1
ON gene_id = mix1.id
JOIN (SELECT id, attamoles_ul AS mix2 FROM ExFold WHERE mix=2) AS mix2
ON gene_id = mix2.id')

factors <-dbGetQuery(mydb,'SELECT * FROM factors')
rownames(factors) <- factors$id
factors$days <- factor(factors$days)
factors$sex <- factor(factors$sex)

mix1<-melt(ercc[,c("gene_id","mix1",factors[factors$mix==1,"id"])],id=c("gene_id","mix1"))
mix2<-melt(ercc[,c("gene_id","mix2",factors[factors$mix==2,"id"])],id=c("gene_id","mix2"))
ggplot(mix2,aes(log(mix2,2),log(value,2),color=variable))+geom_point()
ggplot(mix1,aes(log(mix1,2),log(value,2),color=variable))+geom_point()

```

```{r dge}
fpkm <-dbGetQuery(mydb,'SELECT * FROM fpkm WHERE NOT gene_id REGEXP "^ERCC"')
rownames(fpkm) <- fpkm$gene_id  ; fpkm <- fpkm[,-1]
counts <-dbGetQuery(mydb,'SELECT * FROM counts WHERE NOT gene_id REGEXP "^ERCC"')
rownames(counts) <- counts$gene_id  ; counts <- counts[,-1]
proteins <- dbGetQuery(mydb,'SELECT proteins.*, class FROM proteins JOIN classes ON classes.id = proteins.id GROUP BY proteins.id')
rownames(proteins) <- proteins$id
#blast table previously result from fpkm filtering (>1)
blast <- dbGetQuery(mydb,'SELECT blast.*, rate FROM blast LEFT OUTER JOIN (SELECT id, AVG(rate) AS rate FROM dnds GROUP BY id) dnds1 ON dnds1.id = blast.id')
rownames(blast) <- blast$id 

keep <- (rowMeans(fpkm)>=1 & (rownames(fpkm) %in% rownames(blast)))  #eliminate genes with average fpkm >= 1
table(keep)
#design matrix

design <- model.matrix(~factor(days)+factor(sex)+0,factors)
rownames(design) <- rownames(factors)

#glm
dge <- DGEList(counts[keep,rownames(factors)])
dge <- calcNormFactors(dge)
dat <- estimateGLMCommonDisp(dge, design)
dat <- estimateGLMTrendedDisp(dat, design)
dat <- estimateGLMTagwiseDisp(dat, design)
fit <- glmFit(dat,design)

mds <- plotMDS(dge)
ggpairs(factors[,c("svLength", "tailLength", "totLength", "headLength",  "mass")])
mdsPlot <- data.frame(mds1=mds$cmdscale.out[,1], mds2=mds$cmdscale.out[,2], ID=rownames(mds$cmdscale.out),days=factors$days,sex=factors$sex,mass=factors$mass)
ggplot(mdsPlot,aes(mds1,mds2,shape=sex,color=factor(days),size=log(mass)))+geom_point()+my_theme()

oneVs8 <- glmLRT(fit,contrast=c(1,0,0,-1,0))  
summary(caste.de <- decideTestsDGE(oneVs8, p=0.05, adjust="BH"))
twoVs8 <- glmLRT(fit,contrast=c(0,1,0,-1,0))  
summary(caste.de <- decideTestsDGE(twoVs8, p=0.05, adjust="BH"))
fourVs8 <- glmLRT(fit,contrast=c(0,0,1,-1,0))  
summary(caste.de <- decideTestsDGE(fourVs8, p=0.05, adjust="BH"))
maleVsfemale <- glmLRT(fit,coef=5)  
summary(caste.de <- decideTestsDGE(maleVsfemale, p=0.05, adjust="BH"))
#is there sex-biased expression in venom proteins?
table(p.adjust(maleVsfemale$table[proteins$id,"PValue"],method="fdr")<0.05)
#nope, which makes sense given that the both eat the same things
```


## Filter invariant probes, and plot scale free topology

```{r wgcna_soft_threshold, cache=TRUE}

variance_filter <- function(data) {
  # take WGNCA-style expression as input, and filter it for non-zero variance
  variancedatExpr=as.vector(apply(as.matrix(data),2,var, na.rm=T))
  return(data[, variancedatExpr>0 ])
}

soft_threshold <- function(data) {
  disableWGCNAThreads()
  # Choose a set of soft-thresholding powers, given a WGCNA data object
  powers = seq(from = 14, to=20, by=2)
  # Call the network topology analysis function
  sft = pickSoftThreshold(data, powerVector = powers, verbose = 5,networkType = "signed")
  # Plot the results:
  sizeGrWindow(9, 5)
  par(mfrow = c(1,2));
  cex1 = 0.9;
  # Scale-free topology fit index as a function of the soft-thresholding power
  plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
       xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
      main = paste("Scale independence"));
  text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
      labels=powers,cex=cex1,col="red");
  # this line corresponds to using an R^2 cut-off of h
  abline(h=0.90,col="red")
  # Mean connectivity as a function of the soft-thresholding power
  plot(sft$fitIndices[,1], sft$fitIndices[,5],
      xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
      main = paste("Mean connectivity"))
  text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
  par(mfrow = c(1,1));
    
}
```

```{r setup_wgcna}
se <- DESeqDataSetFromMatrix(round(fpkm[keep,]),factors[,c("days","sex")],formula(~days+sex))
vst <-varianceStabilizingTransformation(se)
# vst <- rlog(se,fitType='local') # not as good
# filter data and plot soft threshold parameters
soft_threshold(variance_filter(t(assay(vst))))

```

Choosing power 16

```{r WGCNA}
enableWGCNAThreads()
wgcna_adjacency <- function(datExpr, threshold, minModuleSize=30, MEDissThres = 0.2, deepSplit = 2) {
  # takes WGNCA data expression matrix, a soft threshold
  # optionally minimum module size, module eigengene dissimilarity threshold, and splitting threthold
  # returns adjacency matrix, module eigengenes and a data frame of colors 
  
  #compute adjacency
  adjacency <- adjacency(datExpr, power = threshold,type="signed")
#  TOM <- TOMsimilarityFromExpr(datExpr,power = threshold, networkType="signed")
  TOM <- TOMsimilarity(adjacency,TOMType="signed")
  
  geneTree <- flashClust(as.dist(1-TOM), method = "average");
  # Module identification using dynamic tree cut:
  dynamicMods <- cutreeDynamic(dendro = geneTree, distM = 1-TOM, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize);
  table(dynamicMods)
  dynamicColors = labels2colors(dynamicMods)
  
  # Calculate eigengenes
  MEList = moduleEigengenes(datExpr, colors = dynamicColors)
  MEs = MEList$eigengenes
  # Calculate dissimilarity of module eigengenes
  METree = flashClust(as.dist(1-cor(MEs)), method = "average");
  plot(METree, main = "Clustering of module eigengenes",xlab = "", sub = "")
  # Plot the cut line into the dendrogram
  abline(h=MEDissThres, col = "red")
  
  merge <- mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 0)
  
  # The merged module colors
  mergedColors = merge$colors
  # Eigengenes of the new merged modules:
  mergedMEs = merge$newMEs
  
  # Rename to moduleColors
  moduleColors = mergedColors
  # Construct numerical labels corresponding to the colors
  colorOrder = c("grey", standardColors(50));
  moduleLabels = match(moduleColors, colorOrder)-1;
  MEs = mergedMEs;
  
  # Recalculate MEs with color labels
  invisible(MEs0 <- moduleEigengenes(datExpr, moduleColors)$eigengenes)
  MEs = orderMEs(MEs0)
  
  print(table(moduleColors))
  moduleColors <- as.data.frame(moduleColors)
  rownames(moduleColors) <- colnames(datExpr)
  
  return(list(adjacency=adjacency,MEs=MEs,moduleColors=moduleColors))
}
```

```{r}
datExpr<-t(assay(vst))
plotClusterTreeSamples(datExpr=datExpr)
results <- wgcna_adjacency(datExpr,16,MEDissThres = 0.25)
#results[["adjacency"]]
#results[["MEs"]]

# look at distribution of proteins in modules
table(results[["moduleColors"]])
table(results[["moduleColors"]][proteins$id,])
proteinsInModules <-data.frame(colors=results[["moduleColors"]][proteins$id,],class=proteins$class,description=proteins$description)
rownames(proteinsInModules)<-proteins$id
table(results[["moduleColors"]][proteins$id,])/table(results[["moduleColors"]])*100
connectivity <- intramodularConnectivity(results[["adjacency"]],results[["moduleColors"]]$moduleColors)

#which proteins change in level across the time series?
for (i in proteins$id) {
  a<-cor.test(as.numeric(factors$days),as.numeric(fpkm[i,factors$id]),method="spearman")
  print(paste(i,proteinsInModules[i,"description"])); print(a$estimate); print (a$p.value)

  plot(as.numeric(factors$days),log(as.numeric(fpkm["c22249_g1",factors$id])))

}
green <- results[["adjacency"]][results[["moduleColors"]]$moduleColor=="green",results[["moduleColors"]]$moduleColor=="green"]
greenAttrib <- data.frame(class=proteins[rownames(green),"class"],blast=blast[rownames(green),"description"])
rownames(greenAttrib) <- rownames(green)
greenAttrib$expressed <- 0
greenAttrib[proteins$id[proteins$id %in% rownames(green)],"expressed"] <- 1
#exportNetworkToVisANT(results[["adjacency"]][results[["moduleColors"]]$moduleColor=="green",results[["moduleColors"]]$moduleColor=="green"],"~/Desktop/mucrosquamatus/adjacency.txt")
exportNetworkToCytoscape(green,edgeFile="~/Desktop/mucrosquamatus/edges.txt",nodeFile="~/Desktop/mucrosquamatus/nodes.txt",nodeAttr=greenAttrib[,"expressed"],nodeNames=rownames(green),includeColNames=TRUE,threshold = 0.5)

```



```{r module_trait}
plotModuleTrait <-function(moduleTraitCor,moduleTraitPvalue,species) {
  # plot heatmap with module-trait correlation data
  textMatrix = paste(signif(moduleTraitCor, 2), " (",
  signif(moduleTraitPvalue, 1), ")", sep = "");
  dim(textMatrix) = dim(moduleTraitCor)
  par(mar = c(6, 8.5, 3, 3));
  # Display the correlation values within a heatmap plot
  labeledHeatmap(Matrix = moduleTraitCor,
  xLabels = colnames(moduleTraitCor),
  yLabels = rownames(moduleTraitCor),
  colorLabels = FALSE,
  colors = blueWhiteRed(50),
  textMatrix = textMatrix,
  setStdMargins = FALSE,
  cex.text = .5,
  zlim = c(-1,1),
  main = paste("Module-trait relationships",species))
}

moduleTraitCor = cor(results[["MEs"]], as.numeric(factors$days), use = "p",method="spearman")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples=30)
print(table(moduleTraitPvalue < 0.05))

ggplot(data.frame(cor=moduleTraitCor,name=rownames(moduleTraitCor),pval=moduleTraitPvalue),aes(x=1,y=name,label=paste(signif(moduleTraitCor, 2), " (", signif(moduleTraitPvalue, 2), ")", sep = "")))+ geom_tile(aes(fill=cor))+ scale_fill_gradient(low = "steelblue",high = "#FF4D4D")+theme_bw()+geom_text()+xlab("")+ylab("")+theme(axis.text.x  = element_blank(),axis.ticks.x  = element_blank())

```


```{r go_term_enricment}
go <- dbGetQuery(mydb,'SELECT go, "ISS" as evidence, id FROM go')
goFrame=GOFrame(go,organism="Protobothrops mucrosquamatus")
universe<-unique(go$id)
goAllFrame=GOAllFrame(goFrame)
gsc <- GeneSetCollection(goAllFrame, setType = GOCollection())

# GO terms upregulated in tan module

thistle2Upreg <- hyperGTest(GSEAGOHyperGParams(name = "thistle2 upregulated", geneSetCollection=gsc, geneIds = intersect(universe,rownames(results[["moduleColors"]])[results[["moduleColors"]]$moduleColors=="thistle2"]), universeGeneIds=universe, ontology = "BP",pvalueCutoff = 0.05,conditional = TRUE,testDirection = "over"))
summary(thistle2Upreg)

greenUpreg <- hyperGTest(GSEAGOHyperGParams(name = "green upregulated", geneSetCollection=gsc, geneIds = intersect(universe,rownames(results[["moduleColors"]])[results[["moduleColors"]]$moduleColors=="green"]), universeGeneIds=universe, ontology = "BP",pvalueCutoff = 0.05,conditional = TRUE,testDirection = "over"))
summary(greenUpreg)

```

```{r exploring_green_module}
green <- data.frame(connectivity = connectivity[results[["moduleColors"]]=="green","kWithin"])
rownames(green) <- rownames(connectivity[results[["moduleColors"]]=="green",])
green$venom <- 0; green[proteins$id,"venom"] <- 1
ggplot(green,aes(connectivity))+geom_histogram()+geom_rug(data=subset(green,venom==1),aes(connectivity,color="red"))+my_theme()
```

```{r exploring_thistle2_module}
thistle2 <- data.frame(connectivity = connectivity[results[["moduleColors"]]=="thistle2","kWithin"])
rownames(thistle2) <- rownames(connectivity[results[["moduleColors"]]=="thistle2",])
thistle2$venom <- 0; thistle2[proteins$id,"venom"] <- 1
ggplot(thistle2,aes(connectivity))+geom_histogram()+geom_rug(data=subset(thistle2,venom==1),aes(connectivity,color="red"))+my_theme()
```

```{r evolutionary_rate_analysis}
rateAnalysis <- subset(na.omit(data.frame(module = results[["moduleColors"]]$moduleColors, rate=blast[rownames(results[["moduleColors"]]),"rate"])),rate<3 & module != "grey")
moduleMeans <- aggregate(rateAnalysis$rate,by=list(rateAnalysis$module),FUN=mean)
rateAnalysis$module <- factor(rateAnalysis$module, levels = moduleMeans[order(-moduleMeans[,2]),1])
ggplot(rateAnalysis,aes(module,rate))+geom_boxplot()+my_theme()
```


```{r connectivity_analysis}
connectivityAnalysis <- data.frame(module = results[["moduleColors"]]$moduleColors, connectivity=connectivity$kWithin)
moduleMeans <- aggregate(connectivityAnalysis$connectivity,by=list(connectivityAnalysis$module),FUN=mean)
connectivityAnalysis$module <- factor(connectivityAnalysis$module, levels = moduleMeans[order(-moduleMeans[,2]),1])

ggplot(connectivityAnalysis,aes(module,connectivity))+geom_boxplot()+my_theme()
connectivityExpressed <- connectivity[rownames(proteins),]
connectivityExpressed$class <- proteins$class
connectivityExpressed$module <- results[["moduleColors"]][rownames(proteins),"moduleColors"]
connectivityExpressed$expression <- rowMeans(fpkm)[rownames(proteins)]
connectivityExpressed$dnds <- blast[rownames(proteins),"rate"]
connectivityExpressed$dnds[connectivityExpressed$dnds > 5] <- NA
ggplot(subset(connectivityExpressed,class!=""),aes(factor(class),kWithin,size=log(expression),color=dnds))+geom_jitter()+my_theme()+theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_grid(.~module,scales="free_x",space="free_x")

connectivity.glm<-glm(sqrt(dnds)~log(expression)*log(kWithin),data=connectivityExpressed)
with(connectivityExpressed,cor.test(dnds,expression,method="spearman"))
hist(resid(connectivity.glm))
summary(connectivity.glm)
ggplot(na.omit(connectivityExpressed),aes(expression,kWithin,color=dnds))+geom_point(size=3)+my_theme()+scale_x_log10()+scale_y_log10()+stat_smooth(method="lm",se=FALSE)
```
